---
title: "Modeling psychological impacts on epidemic spread"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
 
 
# Definitions 
 
 We will try to incorporate opinion dynamics, network modeling, and a bit of the game theory to understand epidemic spread.
 
 
 ## The agent
 
We will use a simplified task network model to represent the biological progression of the disease.
 First, let's suppose that an agent has two states: its psychological state and its biological state.   psychological state might be 'practicing distancing', 'believes conspiracy theory', 'in quarantine', and we can explore these later.  Let's just consider everyone is in a generic 'informed' state [1]. 
 
 The biological state has a few specific cases:
 

1. Unexposed
2. Asymptomatic but infected/contagious
3. Symptomatic and contagious
4. Symptomatic and not contagious
5. Post-COVID Immune
6. Naturally immune (will not contract)
7. Death


We could identify several others, but this might be fine.  So, we can define the agent according to just two numbers.  We might assume that initially, most people are in bio-state 1,  but some would be in state 6 already, which is essentially the same as state 5.

We will just define these as a named list.
```{r}
library(ggplot2)
library(dplyr)
library(sna)
library(knitr)
library(reshape2)
library(igraph)
#library(animation)
STATES <<- 7

STATENAMES <<-  c("Unexposed",
                  "Asymptomatic & contagious",
"Symptomatic and contagious",
"Symptomatic and not contagious",
"Post-COVID immune",
"Naturally immune",
"Death")


STATELABELS <<-  c("Unexposed","Asymptomatic\n & contagious",
"Symptomatic \n& contagious",
"Symptomatic \n& not contagious",
"Post-COVID immune",
"Naturally immune",
"Death")


makeAgent <- function(psych,bio)
{
  return (list(psychstate=psych,biostate=bio))
}
print(makeAgent(1,2))


```


## Timecourse and Biological model

We need a way to transition the biological state in a reasonable way, much like a task network model that keeps track of time of sub-events.  This should consider ONLY the natural progression of the disease. We will model timecourse on a timecourse of 1-day units.


An easy way to do this is with ballistic events.  That is, we can keep track of the next transition point for any state, if it is programmed at the beginning.  This is similar to how I implemented this in the task network model, but a bit simpler because we only have to worry about changes each day.




```{r}


makeAgent <- function(psychstate,biostate,age=30)
{
  
  return (list(psychstate=psychstate,
               biostate=biostate,
               age=age,
               nextbiostate=NA,
               biostatecountdown=NA))
}
print(makeAgent(1,2))


```

This way, we can model the time it takes between becoming infected and showing symptoms.
```{r}
agent <- makeAgent(psychstate=1, biostate=2)

##infect the agent
agent$biostate <- 1   ##infected but asymptomatic
agent$nextbiostate <- 2  ##infected and symptomitac/contageous
agent$biostatecountdown <- ceiling(abs(rnorm(1,mean=8,sd=1)))
print(agent)
```


Now, although the agent is in state 1, it 'knows' it will transition to state 2 in ,e.g., 9 steps. We just need a function to update state generically, which will handle any transition we program into the agent.


```{r}

updateAgent<- function(agent)
{
   agent$biostatecountdown <- agent$biostatecountdown -1
  if(agent$biostatecountdown <=0)
  {
    agent$biostate <- agent$nextbiostate
    agent$biostatecountdown <- NA
  }
   return(agent)
}
```


```{r}
agent$biostate <- 1  ##Unexposed
agent$nextbiostate <- 2  ##asymptomatic but infected
agent$biostatecountdown <- ceiling(abs(rnorm(1,mean=8,sd=1)))
print(agent)
continue <- TRUE
while(continue)
{
  print(paste(agent$biostate,agent$biostatecountdown))
  agent <- updateAgent(agent)
   continue <-!is.na(agent$biostatecountdown)
}
             
```


## Create transition matrix.

If on each timestep we could move between states probabilistically, this would be a markov process. We could represent this by a simple transition network. If we represented the possible transitions in a network, that would simplify the logic above.  But we need more like a task network because the time taken to transition matters.   To keep it simple,   Let's make all timing distributions uniform with a min and max parameter for time in each state.  We can program two pathways through the stages, with a couple branch points (death vs recovery; the possibility of recovering after acquiring with no symptoms).  The progression of the disease is completely specified by the data, and a generic update function will automatically progress the agent each day.


```{r}

# * 1. Unexposed
# * 2. Asymptomatic but infected/contagious
# * 3. Symptomatic and contagious
# * 4. Symptomatic and not contagious
# * 5. Post-COVID Immune
# * 6. Naturally immune (will not contract)
# * 7. Death


bioTransition <- matrix(0,STATES,STATES)
bioMin <- matrix(1,STATES)      #state time minimum
bioMax <- matrix(1,STATES)      #state time maximum



bioMin[2] <- 3            #infected but asymptomatic for 3 to 10 days
bioMax[2] <- 10          
bioTransition[2,3] <- .5  #transition to infected with symptoms
bioTransition[2,5] <- .5  #transition to no longer contagious/cured


bioMin[3] <-    3             #symptoms + contagion
bioMax[3] <- 8                #symptoms + contagion max
bioTransition[3,4] <-  .95    #transitioon to no longer contagious
bioTransition[3,7] <-  .05    #transitioon to death state 


bioMin[4] <- 1          #symptoms bot no longer contagiious
bioMax[4] <- 7
bioTransition[4,5] <- 1  #Transition to 'immune' cured state.




setAgentState<- function(agent, biostate)
{
  agent$biostate <- biostate
  if(sum(bioTransition[biostate,])>0) # this state transitions to something else.
  {
    ##which state do we go to?
     agent$biostatecountdown <- sample(x=seq(bioMin[biostate],bioMax[biostate]),1) #how long will we state in this state?
     agent$nextbiostate <- sample(1:STATES, prob=bioTransition[agent$biostate,],size=1)
     
  } else{
   agent$biostatecountdown <- NA
   agent$nextbiostate <- NA   ##just so we can tell if the agent is finished.
  }
  return(agent) 
}

transitionAgent<- function(agent)
{
   return(setAgentState(agent,agent$nextbiostate))
}

updateAgent<- function(agent)
{
  if(!is.na(agent$biostatecountdown))
  {
   agent$biostatecountdown <- agent$biostatecountdown -1
    if(agent$biostatecountdown <=0)  ##new state
    {
       agent <- transitionAgent(agent)

    }
  }
   return(agent)
}
```

```{r, fig.width=8,fig.height=8}

par(xpd=NA,mar=c(5,5,8,5))
gplot(bioTransition,label=STATELABELS)
```

If we start an agent in an unexposed state, nothing happens.  Agent's do not just naturally develop the disease.
We can just update the agent until there is no next-state programmed.

```{r}

agent <- makeAgent(psychstate=1,biostate=1)

continue <- TRUE
while(continue)
{
  print(paste(agent$biostate,STATENAMES[agent$biostate],agent$biostatecountdown))
  agent <- updateAgent(agent)
  continue <-!is.na(agent$nextbiostate)
}

```





## Social model

This is not quite right, because we'd really expect interactions to be clumpy--you interact with the same people every day, but this model is like the opinion dynamics model in that anyone can interact with anyone.  Let's make a network using preferential attachment to represent that social world.  We can make a couple of them and add them together to represent how we each have different types of relationships and interactions--you may be the president of your drama club and so are a central member, but you are also in a pottery class where you don't talk to very many people (although your teacher does).

```{r,fig.width=20,fig.height=20}


makeNetwork<- function(numAgents,numsets=3,steps=1,power=1)
{
  ord <- sample(numAgents)
  tmp<-as_adjacency_matrix(sample_pa(numAgents,power=power),sparse=F)
  tmp <- tmp + t(tmp)
  tmp <- tmp[ord,ord]
  if(numsets>1)
  {
    for(i in 2:numsets)
  {
    ord <- sample(numAgents)
    sn2 <-as_adjacency_matrix(sample_pa(numAgents,power=power),sparse=F)[ord,ord]
    tmp <- tmp + sn2 + t(sn2)
    
   }
  }
  if(steps>1)
  {
   for(i in 2:steps)
   {
      tmp <- tmp + tmp %*% tmp  ## iterate to neighbors
   }
  }
  (tmp>0)+0
}

## This tries to make a network with families, schools, workplaces..
## 
makeDemographicNetwork <- function(numAgents=1000, 
                                   householdsize=2.5,
                                   numSchools=4,
                                   numWorkplaces=25,
                                   agePyramid = c(20,8,10,12,14,13,12,11))
  
{
  
  numChildren <- floor(sum(agePyramid[1:2])/sum(agePyramid) * numAgents)
  numAdults <- numAgents-numChildren

  ##Adult ages age:
  
  adultAge<- sample(c(3:8)*10-5,size=numAdults,replace=T,prob=agePyramid[-(1:2)])
  childrenAge <- sample(c(5,15),size=numChildren,replace=T,prob=agePyramid[1:2])
  ages <- c(adultAge,childrenAge)
  ##Each household
  numHouseholds <-floor( numAgents/householdsize)
  householdXY <- matrix(runif(numHouseholds*2)*10,ncol=2)
  
  #Now, let's assign people to households.  Adults can be assigned to anywhere;
  ##children need to be assigned to a household that already exists.
  
  ##assign 'head-of-household' ##first the adults, then the children
  myHousehold <- rep(NA,numAgents)
  myHousehold[1:numHouseholds] <- 1:numHouseholds
  myHousehold[(numHouseholds+1):numAgents] <-  sample(numHouseholds,replace=T,
                                                     size=numAgents-numHouseholds)
  
  
  ##now, all households are assigned.  Let's assign workplaces and schools.  
  ##for simplicity, every adult has a workplace, and every child has a school, even though children
  ## or adults may work from home/too young for school, or be retired,
  ##wokplaces 1 is hosptial
  ## workplace 1+1;numschools are considered schools
   
  ##Workplace size is roughly zipfs-law
  workPlaces <-sample(1:numWorkplaces,size=numAdults,prob=1/(1+1:numWorkplaces),replace=T)
  schoolPlaces <- sample(1+(1:numSchools),size=numChildren,replace=T)
  work <- c(workPlaces,rep(0,numChildren))
  school <- c(rep(0,numAdults),schoolPlaces)
  pool <- list()
  for(i in 1:numAgents)
   {
   tmpAgent <- makeAgent(psychstate=1,
                          biostate=1,
                          age=ages[i])

    tmpAgent$household = myHousehold[i]
    tmpAgent$work = work[i]  ##work or school
    tmpAgent$school=school[i]
    tmpAgent$head <- (i<numHouseholds)
    tmpAgent$xy <- householdXY[myHousehold[i] ]
    pool[[i]] <-tmpAgent
   }
 
  
  ##now, let's lay out the social network
  schoolmat <- (outer(school,school,"==") * outer(school,school,"*"))>0
  workmat <- (outer(work,work,"==") * outer(work,work,"*"))>0
  housemat <- ( outer(myHousehold,myHousehold,"=="))>0
  agentXY <- householdXY[myHousehold,] +  matrix(.15* rnorm(length(myHousehold)*2),ncol=2)
  neighborhood <- as.matrix(dist(agentXY)) < runif(nrow(agentXY)^2)*2
#  neighborhood <- outer(agentXY,agentXY, function(a,b){(a-b)^2 < runif(1)*.5})
  network <- schoolmat+ workmat + housemat + neighborhood
   return (list(xy=agentXY, pool=pool,network=network,
                neighborhood=neighborhood,
                worknet=workmat,
                schoolnet=schoolmat,
                housenet=housemat))
}




mygplot <- function(coord, network,states,main="",edgecol="grey40",add=F)
{
  if(is.null(coord))
  {
    coord  <- gplot.layout.fruchtermanreingold(network,layout.par=list(niter=500))
  }
  
  newmin <- mean(coord[,2]) - (-min(coord[,2]) + mean(coord[,2])) * 1.4
   palette=c("white","yellow","red","green","darkgreen","blue","black")
   if(add==F)
     plot(coord,col="black",bty="n",pch=16,cex=2.7,xaxt="n",yaxt="n",main=main,
        xlab="",ylab="",axes=F,
        ylim=c(newmin,max(coord[,2])),type="n")
   
  for(i in 1:nrow(network))
  {
    if(sum(network[i,])>0)
    {
     segments(coord[i,1],
            coord[i,2],
            coord[network[i,]>0,1,drop=F],
            coord[network[i,]>0,2,drop=F],col=edgecol)
    }
   }
       points(coord,pch=16,cex=2.3,col= palette[states])
              
       points(coord,pch=1,cex=2.3,col="black")
      legend(mean(coord[,1]),min(coord[,2]),bty='n',y.intersp=.7,cex=.8,
              STATENAMES, pch=16,col=palette)
              
      return (coord)
}

net <- makeDemographicNetwork(500, numSchools=4,
                                   numWorkplaces=50)

cc <-  mygplot(coord=net$xy,net$housenet,rep(1,nrow(net$network)),main="Households",edgecol="blue")
cc <-  mygplot(coord=net$xy,net$schoolnet+0,rep(1,nrow(net$network)),main="Schools",edgecol="orange")
cc <-  mygplot(coord=net$xy,net$worknet+0,rep(1,nrow(net$network)),main="Workplaces",edgecol="darkgreen")
cc <-  mygplot(coord=net$xy,net$neighborhood,rep(1,nrow(net$network)),
                                    main="Neighborhood network",edgecol="brown")
cc <-  mygplot(coord=net$xy,net$network,rep(1,nrow(net$network)),main="Combined network")
cc <-  mygplot(coord=net$xy,net$housenet,rep(1,nrow(net$network)),main="Household network",edgecol="blue",add=T)
cc <-  mygplot(coord=net$xy,net$schoolnet+0,rep(1,nrow(net$network)),main="School network",edgecol="orange",add=T)
cc <-  mygplot(coord=net$xy,net$worknet+0,rep(1,nrow(net$network)),main="Workplace network",edgecol="darkgreen",add=T)
cc <-  mygplot(coord=net$xy,net$neighborhood,rep(1,nrow(net$network)),
                                    main="Neighborhood network",edgecol="brown",add=T)
```


###No school from day 4 to day 20--90 days:


```{r,fig.width=20,fig.height=20}

numAgents <- 500
numDays <- 90
naturalImmunity <- .01
net <- makeDemographicNetwork(numAgents)

pool <- net$pool
cc <- net$xy
socialnetwork <- net$network


numInteractions <-  rep(20,numDays)  ##how many interactions per day per agent on average?
contagionProb <- rep(.05,numDays)    ##normal contagioun probability after concat
sampleFromNetwork <- rep(.9,numDays)  ##how likely you are to stick with 'your' network

numInteractions[4:numDays] <- 15     ##quarantiine goes into effect day 4.
numInteractions[21:numDays] <- 20  ##re-open 

plotNetwork <- TRUE
#re-use previous network
#socialnetwork <-makeNetwork(numAgents,numsets=1,power=.5)


if(plotNetwork)
{
    cc <-mygplot(coord=cc,socialnetwork,rep(1,nrow(socialnetwork)),main="Initial state")

}



disthistory <- matrix(NA,ncol=7,nrow=numDays)

pool <- list()
for(i in 1:numAgents)
{
   pool[[i]] <- makeAgent(psychstate=1,
                         biostate=sample(c(1,6),
                          p=c(1-naturalImmunity, naturalImmunity),1))
}

##infect patient 0
numInfected <- 5
for(i in sample(numAgents,numInfected))
{
   pool[[i]] <- setAgentState(pool[[i]],2) ##infect this person
}

for(day in 1:numDays)
{

 
    # Create network from individual workplaces/schools/etc. 
  socialnetwork <- net$housenet + net$neighborhood+ net$worknet + net$schoolnet
##who are you going to talk to today.
  if(day>3){
    socialnetwork <- net$housenet+net$neighborhood+net$worknet
  }
  if(day>20){
    socialnetwork <- net$housenet+net$neighborhood +net$worknet + net$schoolnet
  }
    sneezers <- rep(1:numAgents,each=numInteractions[day])
    sneezedons <- rep(NA,length(sneezers))
    
    for(i in 1:length(sneezers))
      {
      if(runif(1)<(1-sampleFromNetwork[day]) )
      {
        sneezedons[i] <- sample(numAgents,1)
      }else{
        sneezedons[i] <- sample(1:numAgents,prob=socialnetwork[sneezers[i],],1)
      }
     }
      
  for(i in 1:length(sneezers))
  {
    agent1 <- pool[[ sneezers[i] ]]
    agent2 <- pool[[ sneezedons[i] ]]
  
    
    ##this constitutes the rules of infection.
    if((agent1$biostate==2 || agent1$biostate==3 ) & agent2$biostate==1 & runif(1)<contagionProb[day])
    {
         pool[[ sneezedons[i] ]] <- setAgentState(agent2,2)##infect!
    }
      
  }
    ##increment each agent 1-day.
    for(i in 1:numAgents)
    {
      pool[[i]] <- updateAgent(pool[[i]])
    }
    
    states <- sapply(pool,FUN=function(x){x$biostate})
    distrib <- table(factor(states,levels=1:7))
    disthistory[day,] <- distrib
   if(plotNetwork)
   {
     mygplot(cc,socialnetwork,states,main=paste("Day",day))
   }
  
         
}
```



```{r fig.width=8,fig.height=5}
disthist.df <-as.data.frame(disthistory)
colnames(disthist.df) <- STATENAMES
disthist.df$day <- 1:nrow(disthistory)
histlong <- melt(disthist.df,id.vars="day")

ggplot(histlong,aes(x=day,y=value,fill=variable)) + geom_bar(stat="identity",position="stack") + 
  theme_bw()

##make the SIR plot:
sir <- data.frame(day=disthist.df$day,
                  susceptible = disthist.df$Unexposed,
                  infected = disthist.df[,2]+disthist.df[,3],
                  recovered = rowSums(disthist.df[,4:7]))

ggplot(melt(sir,id.vars="day"),aes(x=day,group=variable,y=value,color=variable)) + geom_line() + theme_bw()
ggplot(histlong,aes(x=day,y=value,group=variable,color=variable)) + geom_line() + theme_bw()

```




###Check isolation stage: no school, no neighbor day 4 to day 20
```{r,fig.width=20,fig.height=20}

numAgents <- 500
numDays <- 90
naturalImmunity <- .01
net <- makeDemographicNetwork(numAgents)

pool <- net$pool
cc <- net$xy
socialnetwork <- net$network


numInteractions <-  rep(20,numDays)  ##how many interactions per day per agent on average?
contagionProb <- rep(.05,numDays)    ##normal contagioun probability after concat
sampleFromNetwork <- rep(.9,numDays)  ##how likely you are to stick with 'your' network

numInteractions[4:numDays] <- 13     ##quarantiine goes into effect day 4.
numInteractions[21:numDays] <- 20   ##re-open 

plotNetwork <- TRUE
#re-use previous network
#socialnetwork <-makeNetwork(numAgents,numsets=1,power=.5)


if(plotNetwork)
{
    cc <-mygplot(coord=cc,socialnetwork,rep(1,nrow(socialnetwork)),main="Initial state")

}



disthistory <- matrix(NA,ncol=7,nrow=numDays)

pool <- list()
for(i in 1:numAgents)
{
   pool[[i]] <- makeAgent(psychstate=1,
                         biostate=sample(c(1,6),
                          p=c(1-naturalImmunity, naturalImmunity),1))
}

##infect patient 0
numInfected <- 5
for(i in sample(numAgents,numInfected))
{
   pool[[i]] <- setAgentState(pool[[i]],2) ##infect this person
}

for(day in 1:numDays)
{

    # Create network from individual workplaces/schools/etc. 
  socialnetwork <- net$housenet + net$neighborhood+ net$worknet +net$schoolnet
##who are you going to talk to today.
  if(day>3){
    socialnetwork <- net$housenet+net$worknet
  }
  if(day> 21){
    socialnetwork <- net$housenet + net$neighborhood+ net$worknet +net$schoolnet
  }
  
    sneezers <- rep(1:numAgents,each=numInteractions[day])
    sneezedons <- rep(NA,length(sneezers))
    
    for(i in 1:length(sneezers))
      {
      if(runif(1)<(1-sampleFromNetwork[day]) )
      {
        sneezedons[i] <- sample(numAgents,1)
      }else{
        sneezedons[i] <- sample(1:numAgents,prob=socialnetwork[sneezers[i],],1)
      }
     }
      
  for(i in 1:length(sneezers))
  {
    agent1 <- pool[[ sneezers[i] ]]
    agent2 <- pool[[ sneezedons[i] ]]
  
    
    ##this constitutes the rules of infection.
    if((agent1$biostate==2 || agent1$biostate==3 ) & agent2$biostate==1 & runif(1)<contagionProb[day])
    {
         pool[[ sneezedons[i] ]] <- setAgentState(agent2,2)##infect!
    }
      
  }
    ##increment each agent 1-day.
    for(i in 1:numAgents)
    {
      pool[[i]] <- updateAgent(pool[[i]])
    }
    
    states <- sapply(pool,FUN=function(x){x$biostate})
    distrib <- table(factor(states,levels=1:7))
    disthistory[day,] <- distrib
   if(plotNetwork)
   {
     mygplot(cc,socialnetwork,states,main=paste("Day",day))
   }
  
         
}
```
```{r fig.width=8,fig.height=5}
disthist.df <-as.data.frame(disthistory)
colnames(disthist.df) <- STATENAMES
disthist.df$day <- 1:nrow(disthistory)
histlong <- melt(disthist.df,id.vars="day")

ggplot(histlong,aes(x=day,y=value,fill=variable)) + geom_bar(stat="identity",position="stack") + 
  theme_bw()

##make the SIR plot:
sir <- data.frame(day=disthist.df$day,
                  susceptible = disthist.df$Unexposed,
                  infected = disthist.df[,2]+disthist.df[,3],
                  recovered = rowSums(disthist.df[,4:7]))

ggplot(melt(sir,id.vars="day"),aes(x=day,group=variable,y=value,color=variable)) + geom_line() + theme_bw()
ggplot(histlong,aes(x=day,y=value,group=variable,color=variable)) + geom_line() + theme_bw()
```

###Check isolation stage: no school, no neighbor, no workplace from day 4 to day 20
```{r,fig.width=20,fig.height=20}
#network <- housemat
numAgents <- 500
numDays <- 90
naturalImmunity <- .01
net <- makeDemographicNetwork(numAgents)

pool <- net$pool
cc <- net$xy
socialnetwork <- net$network


numInteractions <-  rep(20,numDays)  ##how many interactions per day per agent on average?
contagionProb <- rep(.05,numDays)    ##normal contagioun probability after concat
sampleFromNetwork <- rep(.9,numDays)  ##how likely you are to stick with 'your' network

numInteractions[4:numDays] <- 3     ##quarantiine goes into effect day 4.
numInteractions[21:numDays] <- 20  ##re-open 

plotNetwork <- TRUE
#re-use previous network
#socialnetwork <-makeNetwork(numAgents,numsets=1,power=.5)


if(plotNetwork)
{
    cc <-mygplot(coord=cc,socialnetwork,rep(1,nrow(socialnetwork)),main="Initial state")

}



disthistory <- matrix(NA,ncol=7,nrow=numDays)

pool <- list()
for(i in 1:numAgents)
{
   pool[[i]] <- makeAgent(psychstate=1,
                         biostate=sample(c(1,6),
                          p=c(1-naturalImmunity, naturalImmunity),1))
}

##infect patient 0
numInfected <- 5
for(i in sample(numAgents,numInfected))
{
   pool[[i]] <- setAgentState(pool[[i]],2) ##infect this person
}

for(day in 1:numDays)
{

   
    # Create network from individual workplaces/schools/etc. 
  socialnetwork <- net$housenet + net$neighborhood+ net$worknet +net$schoolnet
##who are you going to talk to today.
  if(day>3 ){
    socialnetwork <- net$housenet
  }
  if(day>21 ){
    socialnetwork <- net$housenet + net$neighborhood+ net$worknet +net$schoolnet
  }
    sneezers <- rep(1:numAgents,each=numInteractions[day])
    sneezedons <- rep(NA,length(sneezers))
    
    for(i in 1:length(sneezers))
      {
      if(runif(1)<(1-sampleFromNetwork[day]) )
      {
        sneezedons[i] <- sample(numAgents,1)
      }else{
        sneezedons[i] <- sample(1:numAgents,prob=socialnetwork[sneezers[i],],1)
      }
     }
      
  for(i in 1:length(sneezers))
  {
    agent1 <- pool[[ sneezers[i] ]]
    agent2 <- pool[[ sneezedons[i] ]]
  
    
    ##this constitutes the rules of infection.
    if((agent1$biostate==2 || agent1$biostate==3 ) & agent2$biostate==1 & runif(1)<contagionProb[day])
    {
         pool[[ sneezedons[i] ]] <- setAgentState(agent2,2)##infect!
    }
      
  }
    ##increment each agent 1-day.
    for(i in 1:numAgents)
    {
      pool[[i]] <- updateAgent(pool[[i]])
    }
    
    states <- sapply(pool,FUN=function(x){x$biostate})
    distrib <- table(factor(states,levels=1:7))
    disthistory[day,] <- distrib
   if(plotNetwork)
   {
     mygplot(cc,socialnetwork,states,main=paste("Day",day))
   }
  
         
}
```
```{r fig.width=8,fig.height=5}
disthist.df <-as.data.frame(disthistory)
colnames(disthist.df) <- STATENAMES
disthist.df$day <- 1:nrow(disthistory)
histlong <- melt(disthist.df,id.vars="day")

ggplot(histlong,aes(x=day,y=value,fill=variable)) + geom_bar(stat="identity",position="stack") + 
  theme_bw()

##make the SIR plot:
sir <- data.frame(day=disthist.df$day,
                  susceptible = disthist.df$Unexposed,
                  infected = disthist.df[,2]+disthist.df[,3],
                  recovered = rowSums(disthist.df[,4:7]))

ggplot(melt(sir,id.vars="day"),aes(x=day,group=variable,y=value,color=variable)) + geom_line() + theme_bw()
ggplot(histlong,aes(x=day,y=value,group=variable,color=variable)) + geom_line() + theme_bw()
```

###Check isolation stage: no school, no workplace from day 4 to day 21
```{r,fig.width=20,fig.height=20}
numAgents <- 500
numDays <- 90
naturalImmunity <- .01
net <- makeDemographicNetwork(numAgents)

pool <- net$pool
cc <- net$xy
socialnetwork <- net$network


numInteractions <-  rep(20,numDays)  ##how many interactions per day per agent on average?
contagionProb <- rep(.05,numDays)    ##normal contagioun probability after concat
sampleFromNetwork <- rep(.9,numDays)  ##how likely you are to stick with 'your' network

numInteractions[4:numDays] <- 5    ##quarantiine goes into effect day 4.
numInteractions[21:numDays] <- 20  ##re-open 

plotNetwork <- TRUE
#re-use previous network
#socialnetwork <-makeNetwork(numAgents,numsets=1,power=.5)


if(plotNetwork)
{
    cc <-mygplot(coord=cc,socialnetwork,rep(1,nrow(socialnetwork)),main="Initial state")

}



disthistory <- matrix(NA,ncol=7,nrow=numDays)

pool <- list()
for(i in 1:numAgents)
{
   pool[[i]] <- makeAgent(psychstate=1,
                         biostate=sample(c(1,6),
                          p=c(1-naturalImmunity, naturalImmunity),1))
}

##infect patient 0
numInfected <- 5
for(i in sample(numAgents,numInfected))
{
   pool[[i]] <- setAgentState(pool[[i]],2) ##infect this person
}

for(day in 1:numDays)
{

  
    # Create network from individual workplaces/schools/etc. 
  socialnetwork <- net$housenet + net$neighborhood+ net$worknet +net$schoolnet
##who are you going to talk to today.
  if(day>3){
    socialnetwork <- net$housenet+net$neighborhood
  }
  if(day> 21){
    socialnetwork <- net$housenet+net$neighborhood+ net$worknet +net$schoolnet
  }
    sneezers <- rep(1:numAgents,each=numInteractions[day])
    sneezedons <- rep(NA,length(sneezers))
    
    for(i in 1:length(sneezers))
      {
      if(runif(1)<(1-sampleFromNetwork[day]) )
      {
        sneezedons[i] <- sample(numAgents,1)
      }else{
        sneezedons[i] <- sample(1:numAgents,prob=socialnetwork[sneezers[i],],1)
      }
     }
      
  for(i in 1:length(sneezers))
  {
    agent1 <- pool[[ sneezers[i] ]]
    agent2 <- pool[[ sneezedons[i] ]]
  
    
    ##this constitutes the rules of infection.
    if((agent1$biostate==2 || agent1$biostate==3 ) & agent2$biostate==1 & runif(1)<contagionProb[day])
    {
         pool[[ sneezedons[i] ]] <- setAgentState(agent2,2)##infect!
    }
      
  }
    ##increment each agent 1-day.
    for(i in 1:numAgents)
    {
      pool[[i]] <- updateAgent(pool[[i]])
    }
    
    states <- sapply(pool,FUN=function(x){x$biostate})
    distrib <- table(factor(states,levels=1:7))
    disthistory[day,] <- distrib
   if(plotNetwork)
   {
     mygplot(cc,socialnetwork,states,main=paste("Day",day))
   }
  
         
}
```
```{r fig.width=8,fig.height=5}
disthist.df <-as.data.frame(disthistory)
colnames(disthist.df) <- STATENAMES
disthist.df$day <- 1:nrow(disthistory)
histlong <- melt(disthist.df,id.vars="day")

ggplot(histlong,aes(x=day,y=value,fill=variable)) + geom_bar(stat="identity",position="stack") + 
  theme_bw()

##make the SIR plot:
sir <- data.frame(day=disthist.df$day,
                  susceptible = disthist.df$Unexposed,
                  infected = disthist.df[,2]+disthist.df[,3],
                  recovered = rowSums(disthist.df[,4:7]))

ggplot(melt(sir,id.vars="day"),aes(x=day,group=variable,y=value,color=variable)) + geom_line() + theme_bw()
ggplot(histlong,aes(x=day,y=value,group=variable,color=variable)) + geom_line() + theme_bw()
```

###Check isolation stage: no school for rest of the year but  no workplace for only 20 days
```{r,fig.width=20,fig.height=20}
#network <- housemat
numAgents <- 500
numDays <- 90
naturalImmunity <- .01
net <- makeDemographicNetwork(numAgents)

pool <- net$pool
cc <- net$xy
socialnetwork <- net$network


numInteractions <-  rep(20,numDays)  ##how many interactions per day per agent on average?
contagionProb <- rep(.05,numDays)    ##normal contagioun probability after concat
sampleFromNetwork <- rep(.9,numDays)  ##how likely you are to stick with 'your' network

numInteractions[4:numDays] <- 5     ##quarantiine goes into effect day 3.
numInteractions[21:numDays] <- 15  ##re-open 

plotNetwork <- TRUE
#re-use previous network
#socialnetwork <-makeNetwork(numAgents,numsets=1,power=.5)


if(plotNetwork)
{
    cc <-mygplot(coord=cc,socialnetwork,rep(1,nrow(socialnetwork)),main="Initial state")

}



disthistory <- matrix(NA,ncol=7,nrow=numDays)

pool <- list()
for(i in 1:numAgents)
{
   pool[[i]] <- makeAgent(psychstate=1,
                         biostate=sample(c(1,6),
                          p=c(1-naturalImmunity, naturalImmunity),1))
}

##infect patient 0
numInfected <- 5
for(i in sample(numAgents,numInfected))
{
   pool[[i]] <- setAgentState(pool[[i]],2) ##infect this person
}

for(day in 1:numDays)
{

  
    # Create network from individual workplaces/schools/etc. 
  socialnetwork <- net$housenet + net$neighborhood+ net$worknet +net$schoolnet
##who are you going to talk to today.
  if(day>3){
    socialnetwork <- net$housenet+net$neighborhood
  }
    if(day>21){
    socialnetwork <- net$housenet+net$neighborhood+net$worknet
  }
    sneezers <- rep(1:numAgents,each=numInteractions[day])
    sneezedons <- rep(NA,length(sneezers))
    
    for(i in 1:length(sneezers))
      {
      if(runif(1)<(1-sampleFromNetwork[day]) )
      {
        sneezedons[i] <- sample(numAgents,1)
      }else{
        sneezedons[i] <- sample(1:numAgents,prob=socialnetwork[sneezers[i],],1)
      }
     }
      
  for(i in 1:length(sneezers))
  {
    agent1 <- pool[[ sneezers[i] ]]
    agent2 <- pool[[ sneezedons[i] ]]
  
    
    ##this constitutes the rules of infection.
    if((agent1$biostate==2 || agent1$biostate==3 ) & agent2$biostate==1 & runif(1)<contagionProb[day])
    {
         pool[[ sneezedons[i] ]] <- setAgentState(agent2,2)##infect!
    }
      
  }
    ##increment each agent 1-day.
    for(i in 1:numAgents)
    {
      pool[[i]] <- updateAgent(pool[[i]])
    }
    
    states <- sapply(pool,FUN=function(x){x$biostate})
    distrib <- table(factor(states,levels=1:7))
    disthistory[day,] <- distrib
   if(plotNetwork)
   {
     mygplot(cc,socialnetwork,states,main=paste("Day",day))
   }
  
         
}
```

```{r fig.width=8,fig.height=5}
disthist.df <-as.data.frame(disthistory)
colnames(disthist.df) <- STATENAMES
disthist.df$day <- 1:nrow(disthistory)
histlong <- melt(disthist.df,id.vars="day")

ggplot(histlong,aes(x=day,y=value,fill=variable)) + geom_bar(stat="identity",position="stack") + 
  theme_bw()

##make the SIR plot:
sir <- data.frame(day=disthist.df$day,
                  susceptible = disthist.df$Unexposed,
                  infected = disthist.df[,2]+disthist.df[,3],
                  recovered = rowSums(disthist.df[,4:7]))

ggplot(melt(sir,id.vars="day"),aes(x=day,group=variable,y=value,color=variable)) + geom_line() + theme_bw()
ggplot(histlong,aes(x=day,y=value,group=variable,color=variable)) + geom_line() + theme_bw()
```

###Check isolation stage: no school but  no workplace from day 4 to day 27 day
```{r,fig.width=20,fig.height=20}

numAgents <- 500
numDays <- 90
naturalImmunity <- .01
net <- makeDemographicNetwork(numAgents)

pool <- net$pool
cc <- net$xy
socialnetwork <- net$network


numInteractions <-  rep(20,numDays)  ##how many interactions per day per agent on average?
contagionProb <- rep(.05,numDays)    ##normal contagioun probability after concat
sampleFromNetwork <- rep(.9,numDays)  ##how likely you are to stick with 'your' network

numInteractions[4:numDays] <- 5     ##quarantiine goes into effect day 3.
numInteractions[27:numDays] <- 15  ##re-open 

plotNetwork <- TRUE
#re-use previous network
#socialnetwork <-makeNetwork(numAgents,numsets=1,power=.5)


if(plotNetwork)
{
    cc <-mygplot(coord=cc,socialnetwork,rep(1,nrow(socialnetwork)),main="Initial state")

}



disthistory <- matrix(NA,ncol=7,nrow=numDays)

pool <- list()
for(i in 1:numAgents)
{
   pool[[i]] <- makeAgent(psychstate=1,
                         biostate=sample(c(1,6),
                          p=c(1-naturalImmunity, naturalImmunity),1))
}

##infect patient 0
numInfected <- 5
for(i in sample(numAgents,numInfected))
{
   pool[[i]] <- setAgentState(pool[[i]],2) ##infect this person
}

for(day in 1:numDays)
{

  # Create network from individual workplaces/schools/etc. 
  socialnetwork <- net$housenet + net$neighborhood+ net$worknet +net$schoolnet
##who are you going to talk to today.
  if(day>3){
    socialnetwork <- net$housenet+net$neighborhood
  }
  if(day>27){
    socialnetwork <- net$housenet+net$neighborhood +net$worknet
  }
    sneezers <- rep(1:numAgents,each=numInteractions[day])
    sneezedons <- rep(NA,length(sneezers))
    
    for(i in 1:length(sneezers))
      {
      if(runif(1)<(1-sampleFromNetwork[day]) )
      {
        sneezedons[i] <- sample(numAgents,1)
      }else{
        sneezedons[i] <- sample(1:numAgents,prob=socialnetwork[sneezers[i],],1)
      }
     }
      
  for(i in 1:length(sneezers))
  {
    agent1 <- pool[[ sneezers[i] ]]
    agent2 <- pool[[ sneezedons[i] ]]
  
    
    ##this constitutes the rules of infection.
    if((agent1$biostate==2 || agent1$biostate==3 ) & agent2$biostate==1 & runif(1)<contagionProb[day])
    {
         pool[[ sneezedons[i] ]] <- setAgentState(agent2,2)##infect!
    }
      
  }
    ##increment each agent 1-day.
    for(i in 1:numAgents)
    {
      pool[[i]] <- updateAgent(pool[[i]])
    }
    
    states <- sapply(pool,FUN=function(x){x$biostate})
    distrib <- table(factor(states,levels=1:7))
    disthistory[day,] <- distrib
   if(plotNetwork)
   {
     mygplot(cc,socialnetwork,states,main=paste("Day",day))
   }
  
         
}
```
```{r fig.width=8,fig.height=5}
disthist.df <-as.data.frame(disthistory)
colnames(disthist.df) <- STATENAMES
disthist.df$day <- 1:nrow(disthistory)
histlong <- melt(disthist.df,id.vars="day")

ggplot(histlong,aes(x=day,y=value,fill=variable)) + geom_bar(stat="identity",position="stack") + 
  theme_bw()

##make the SIR plot:
sir <- data.frame(day=disthist.df$day,
                  susceptible = disthist.df$Unexposed,
                  infected = disthist.df[,2]+disthist.df[,3],
                  recovered = rowSums(disthist.df[,4:7]))

ggplot(melt(sir,id.vars="day"),aes(x=day,group=variable,y=value,color=variable)) + geom_line() + theme_bw()
ggplot(histlong,aes(x=day,y=value,group=variable,color=variable)) + geom_line() + theme_bw()
```





###Check isolation stage: no school but  no workplace from day 4 to day 34 
```{r,fig.width=20,fig.height=20}

numAgents <- 500
numDays <- 90
naturalImmunity <- .01
net <- makeDemographicNetwork(numAgents)

pool <- net$pool
cc <- net$xy
socialnetwork <- net$network


numInteractions <-  rep(20,numDays)  ##how many interactions per day per agent on average?
contagionProb <- rep(.05,numDays)    ##normal contagioun probability after concat
sampleFromNetwork <- rep(.9,numDays)  ##how likely you are to stick with 'your' network

numInteractions[4:numDays] <- 5     ##quarantiine goes into effect day 3.
numInteractions[34:numDays] <- 15  ##re-open 

plotNetwork <- TRUE
#re-use previous network
#socialnetwork <-makeNetwork(numAgents,numsets=1,power=.5)


if(plotNetwork)
{
    cc <-mygplot(coord=cc,socialnetwork,rep(1,nrow(socialnetwork)),main="Initial state")

}



disthistory <- matrix(NA,ncol=7,nrow=numDays)

pool <- list()
for(i in 1:numAgents)
{
   pool[[i]] <- makeAgent(psychstate=1,
                         biostate=sample(c(1,6),
                          p=c(1-naturalImmunity, naturalImmunity),1))
}

##infect patient 0
numInfected <- 5
for(i in sample(numAgents,numInfected))
{
   pool[[i]] <- setAgentState(pool[[i]],2) ##infect this person
}

for(day in 1:numDays)
{

  # Create network from individual workplaces/schools/etc. 
  socialnetwork <- net$housenet + net$neighborhood+ net$worknet +net$schoolnet
##who are you going to talk to today.
  if(day>3){
    socialnetwork <- net$housenet+net$neighborhood
  }
  if(day>34){
    socialnetwork <- net$housenet+net$neighborhood +net$worknet
  }
    sneezers <- rep(1:numAgents,each=numInteractions[day])
    sneezedons <- rep(NA,length(sneezers))
    
    for(i in 1:length(sneezers))
      {
      if(runif(1)<(1-sampleFromNetwork[day]) )
      {
        sneezedons[i] <- sample(numAgents,1)
      }else{
        sneezedons[i] <- sample(1:numAgents,prob=socialnetwork[sneezers[i],],1)
      }
     }
      
  for(i in 1:length(sneezers))
  {
    agent1 <- pool[[ sneezers[i] ]]
    agent2 <- pool[[ sneezedons[i] ]]
  
    
    ##this constitutes the rules of infection.
    if((agent1$biostate==2 || agent1$biostate==3 ) & agent2$biostate==1 & runif(1)<contagionProb[day])
    {
         pool[[ sneezedons[i] ]] <- setAgentState(agent2,2)##infect!
    }
      
  }
    ##increment each agent 1-day.
    for(i in 1:numAgents)
    {
      pool[[i]] <- updateAgent(pool[[i]])
    }
    
    states <- sapply(pool,FUN=function(x){x$biostate})
    distrib <- table(factor(states,levels=1:7))
    disthistory[day,] <- distrib
   if(plotNetwork)
   {
     mygplot(cc,socialnetwork,states,main=paste("Day",day))
   }
  
         
}
```
```{r fig.width=8,fig.height=5}
disthist.df <-as.data.frame(disthistory)
colnames(disthist.df) <- STATENAMES
disthist.df$day <- 1:nrow(disthistory)
histlong <- melt(disthist.df,id.vars="day")

ggplot(histlong,aes(x=day,y=value,fill=variable)) + geom_bar(stat="identity",position="stack") + 
  theme_bw()

##make the SIR plot:
sir <- data.frame(day=disthist.df$day,
                  susceptible = disthist.df$Unexposed,
                  infected = disthist.df[,2]+disthist.df[,3],
                  recovered = rowSums(disthist.df[,4:7]))

ggplot(melt(sir,id.vars="day"),aes(x=day,group=variable,y=value,color=variable)) + geom_line() + theme_bw()
ggplot(histlong,aes(x=day,y=value,group=variable,color=variable)) + geom_line() + theme_bw()
```

###Check isolation stage: no school but  no workplace from day 4 to day 41 
```{r,fig.width=20,fig.height=20}

numAgents <- 500
numDays <- 90
naturalImmunity <- .01
net <- makeDemographicNetwork(numAgents)

pool <- net$pool
cc <- net$xy
socialnetwork <- net$network


numInteractions <-  rep(20,numDays)  ##how many interactions per day per agent on average?
contagionProb <- rep(.05,numDays)    ##normal contagioun probability after concat
sampleFromNetwork <- rep(.9,numDays)  ##how likely you are to stick with 'your' network

numInteractions[4:numDays] <- 5     ##quarantiine goes into effect day 3.
numInteractions[41:numDays] <- 15  ##re-open 

plotNetwork <- TRUE
#re-use previous network
#socialnetwork <-makeNetwork(numAgents,numsets=1,power=.5)


if(plotNetwork)
{
    cc <-mygplot(coord=cc,socialnetwork,rep(1,nrow(socialnetwork)),main="Initial state")

}



disthistory <- matrix(NA,ncol=7,nrow=numDays)

pool <- list()
for(i in 1:numAgents)
{
   pool[[i]] <- makeAgent(psychstate=1,
                         biostate=sample(c(1,6),
                          p=c(1-naturalImmunity, naturalImmunity),1))
}

##infect patient 0
numInfected <- 5
for(i in sample(numAgents,numInfected))
{
   pool[[i]] <- setAgentState(pool[[i]],2) ##infect this person
}

for(day in 1:numDays)
{

  # Create network from individual workplaces/schools/etc. 
  socialnetwork <- net$housenet + net$neighborhood+ net$worknet +net$schoolnet
##who are you going to talk to today.
  if(day>3){
    socialnetwork <- net$housenet+net$neighborhood
  }
  if(day>41){
    socialnetwork <- net$housenet+net$neighborhood +net$worknet
  }
    sneezers <- rep(1:numAgents,each=numInteractions[day])
    sneezedons <- rep(NA,length(sneezers))
    
    for(i in 1:length(sneezers))
      {
      if(runif(1)<(1-sampleFromNetwork[day]) )
      {
        sneezedons[i] <- sample(numAgents,1)
      }else{
        sneezedons[i] <- sample(1:numAgents,prob=socialnetwork[sneezers[i],],1)
      }
     }
      
  for(i in 1:length(sneezers))
  {
    agent1 <- pool[[ sneezers[i] ]]
    agent2 <- pool[[ sneezedons[i] ]]
  
    
    ##this constitutes the rules of infection.
    if((agent1$biostate==2 || agent1$biostate==3 ) & agent2$biostate==1 & runif(1)<contagionProb[day])
    {
         pool[[ sneezedons[i] ]] <- setAgentState(agent2,2)##infect!
    }
      
  }
    ##increment each agent 1-day.
    for(i in 1:numAgents)
    {
      pool[[i]] <- updateAgent(pool[[i]])
    }
    
    states <- sapply(pool,FUN=function(x){x$biostate})
    distrib <- table(factor(states,levels=1:7))
    disthistory[day,] <- distrib
   if(plotNetwork)
   {
     mygplot(cc,socialnetwork,states,main=paste("Day",day))
   }
  
         
}
```

```{r fig.width=8,fig.height=5}
disthist.df <-as.data.frame(disthistory)
colnames(disthist.df) <- STATENAMES
disthist.df$day <- 1:nrow(disthistory)
histlong <- melt(disthist.df,id.vars="day")

ggplot(histlong,aes(x=day,y=value,fill=variable)) + geom_bar(stat="identity",position="stack") + 
  theme_bw()

##make the SIR plot:
sir <- data.frame(day=disthist.df$day,
                  susceptible = disthist.df$Unexposed,
                  infected = disthist.df[,2]+disthist.df[,3],
                  recovered = rowSums(disthist.df[,4:7]))

ggplot(melt(sir,id.vars="day"),aes(x=day,group=variable,y=value,color=variable)) + geom_line() + theme_bw()
ggplot(histlong,aes(x=day,y=value,group=variable,color=variable)) + geom_line() + theme_bw()
```



###Check isolation stage: no school but  no workplace from day 4 to day 48 
```{r,fig.width=20,fig.height=20}

numAgents <- 500
numDays <- 90
naturalImmunity <- .01
net <- makeDemographicNetwork(numAgents)

pool <- net$pool
cc <- net$xy
socialnetwork <- net$network


numInteractions <-  rep(20,numDays)  ##how many interactions per day per agent on average?
contagionProb <- rep(.05,numDays)    ##normal contagioun probability after concat
sampleFromNetwork <- rep(.9,numDays)  ##how likely you are to stick with 'your' network

numInteractions[4:numDays] <- 5     ##quarantiine goes into effect day 3.
numInteractions[48:numDays] <- 15  ##re-open 

plotNetwork <- TRUE
#re-use previous network
#socialnetwork <-makeNetwork(numAgents,numsets=1,power=.5)


if(plotNetwork)
{
    cc <-mygplot(coord=cc,socialnetwork,rep(1,nrow(socialnetwork)),main="Initial state")

}



disthistory <- matrix(NA,ncol=7,nrow=numDays)

pool <- list()
for(i in 1:numAgents)
{
   pool[[i]] <- makeAgent(psychstate=1,
                         biostate=sample(c(1,6),
                          p=c(1-naturalImmunity, naturalImmunity),1))
}

##infect patient 0
numInfected <- 5
for(i in sample(numAgents,numInfected))
{
   pool[[i]] <- setAgentState(pool[[i]],2) ##infect this person
}

for(day in 1:numDays)
{

  # Create network from individual workplaces/schools/etc. 
  socialnetwork <- net$housenet + net$neighborhood+ net$worknet +net$schoolnet
##who are you going to talk to today.
  if(day>3){
    socialnetwork <- net$housenet+net$neighborhood
  }
  if(day>48){
    socialnetwork <- net$housenet+net$neighborhood +net$worknet
  }
    sneezers <- rep(1:numAgents,each=numInteractions[day])
    sneezedons <- rep(NA,length(sneezers))
    
    for(i in 1:length(sneezers))
      {
      if(runif(1)<(1-sampleFromNetwork[day]) )
      {
        sneezedons[i] <- sample(numAgents,1)
      }else{
        sneezedons[i] <- sample(1:numAgents,prob=socialnetwork[sneezers[i],],1)
      }
     }
      
  for(i in 1:length(sneezers))
  {
    agent1 <- pool[[ sneezers[i] ]]
    agent2 <- pool[[ sneezedons[i] ]]
  
    
    ##this constitutes the rules of infection.
    if((agent1$biostate==2 || agent1$biostate==3 ) & agent2$biostate==1 & runif(1)<contagionProb[day])
    {
         pool[[ sneezedons[i] ]] <- setAgentState(agent2,2)##infect!
    }
      
  }
    ##increment each agent 1-day.
    for(i in 1:numAgents)
    {
      pool[[i]] <- updateAgent(pool[[i]])
    }
    
    states <- sapply(pool,FUN=function(x){x$biostate})
    distrib <- table(factor(states,levels=1:7))
    disthistory[day,] <- distrib
   if(plotNetwork)
   {
     mygplot(cc,socialnetwork,states,main=paste("Day",day))
   }
  
         
}
```

```{r fig.width=8,fig.height=5}
disthist.df <-as.data.frame(disthistory)
colnames(disthist.df) <- STATENAMES
disthist.df$day <- 1:nrow(disthistory)
histlong <- melt(disthist.df,id.vars="day")

ggplot(histlong,aes(x=day,y=value,fill=variable)) + geom_bar(stat="identity",position="stack") + 
  theme_bw()

##make the SIR plot:
sir <- data.frame(day=disthist.df$day,
                  susceptible = disthist.df$Unexposed,
                  infected = disthist.df[,2]+disthist.df[,3],
                  recovered = rowSums(disthist.df[,4:7]))

ggplot(melt(sir,id.vars="day"),aes(x=day,group=variable,y=value,color=variable)) + geom_line() + theme_bw()
ggplot(histlong,aes(x=day,y=value,group=variable,color=variable)) + geom_line() + theme_bw()
```


###Check isolation stage: no school but  no workplace from day 4 to day 55 
```{r,fig.width=20,fig.height=20}

numAgents <- 500
numDays <- 90
naturalImmunity <- .01
net <- makeDemographicNetwork(numAgents)

pool <- net$pool
cc <- net$xy
socialnetwork <- net$network


numInteractions <-  rep(20,numDays)  ##how many interactions per day per agent on average?
contagionProb <- rep(.05,numDays)    ##normal contagioun probability after concat
sampleFromNetwork <- rep(.9,numDays)  ##how likely you are to stick with 'your' network

numInteractions[4:numDays] <- 5     ##quarantiine goes into effect day 3.
numInteractions[55:numDays] <- 15  ##re-open 

plotNetwork <- TRUE
#re-use previous network
#socialnetwork <-makeNetwork(numAgents,numsets=1,power=.5)


if(plotNetwork)
{
    cc <-mygplot(coord=cc,socialnetwork,rep(1,nrow(socialnetwork)),main="Initial state")

}



disthistory <- matrix(NA,ncol=7,nrow=numDays)

pool <- list()
for(i in 1:numAgents)
{
   pool[[i]] <- makeAgent(psychstate=1,
                         biostate=sample(c(1,6),
                          p=c(1-naturalImmunity, naturalImmunity),1))
}

##infect patient 0
numInfected <- 5
for(i in sample(numAgents,numInfected))
{
   pool[[i]] <- setAgentState(pool[[i]],2) ##infect this person
}

for(day in 1:numDays)
{

  # Create network from individual workplaces/schools/etc. 
 socialnetwork <- net$housenet + net$neighborhood+ net$worknet +net$schoolnet
##who are you going to talk to today.
  if(day>3){
    socialnetwork <- net$housenet+net$neighborhood
  }
  if(day>55){
    socialnetwork <- net$housenet+net$neighborhood +net$worknet
  }
    sneezers <- rep(1:numAgents,each=numInteractions[day])
    sneezedons <- rep(NA,length(sneezers))
    
    for(i in 1:length(sneezers))
      {
      if(runif(1)<(1-sampleFromNetwork[day]) )
      {
        sneezedons[i] <- sample(numAgents,1)
      }else{
        sneezedons[i] <- sample(1:numAgents,prob=socialnetwork[sneezers[i],],1)
      }
     }
      
  for(i in 1:length(sneezers))
  {
    agent1 <- pool[[ sneezers[i] ]]
    agent2 <- pool[[ sneezedons[i] ]]
  
    
    ##this constitutes the rules of infection.
    if((agent1$biostate==2 || agent1$biostate==3 ) & agent2$biostate==1 & runif(1)<contagionProb[day])
    {
         pool[[ sneezedons[i] ]] <- setAgentState(agent2,2)##infect!
    }
      
  }
    ##increment each agent 1-day.
    for(i in 1:numAgents)
    {
      pool[[i]] <- updateAgent(pool[[i]])
    }
    
    states <- sapply(pool,FUN=function(x){x$biostate})
    distrib <- table(factor(states,levels=1:7))
    disthistory[day,] <- distrib
   if(plotNetwork)
   {
     mygplot(cc,socialnetwork,states,main=paste("Day",day))
   }
  
         
}
```

```{r fig.width=8,fig.height=5}
disthist.df <-as.data.frame(disthistory)
colnames(disthist.df) <- STATENAMES
disthist.df$day <- 1:nrow(disthistory)
histlong <- melt(disthist.df,id.vars="day")

ggplot(histlong,aes(x=day,y=value,fill=variable)) + geom_bar(stat="identity",position="stack") + 
  theme_bw()

##make the SIR plot:
sir <- data.frame(day=disthist.df$day,
                  susceptible = disthist.df$Unexposed,
                  infected = disthist.df[,2]+disthist.df[,3],
                  recovered = rowSums(disthist.df[,4:7]))

ggplot(melt(sir,id.vars="day"),aes(x=day,group=variable,y=value,color=variable)) + geom_line() + theme_bw()
ggplot(histlong,aes(x=day,y=value,group=variable,color=variable)) + geom_line() + theme_bw()
```

----------------------------------------
----------------------------------------

####99% likely to stick with the network



###Check isolation stage: no school for rest of the year but  no workplace for only 20 days
```{r,fig.width=20,fig.height=20}

numAgents <- 500
numDays <- 90
naturalImmunity <- .01
net <- makeDemographicNetwork(numAgents)

pool <- net$pool
cc <- net$xy
socialnetwork <- net$network


numInteractions <-  rep(20,numDays)  ##how many interactions per day per agent on average?
contagionProb <- rep(.05,numDays)    ##normal contagioun probability after concat
sampleFromNetwork <- rep(.99,numDays)  ##how likely you are to stick with 'your' network

numInteractions[4:numDays] <- 5     ##quarantiine goes into effect day 3.
numInteractions[21:numDays] <- 15  ##re-open 

plotNetwork <- TRUE
#re-use previous network
#socialnetwork <-makeNetwork(numAgents,numsets=1,power=.5)


if(plotNetwork)
{
    cc <-mygplot(coord=cc,socialnetwork,rep(1,nrow(socialnetwork)),main="Initial state")

}



disthistory <- matrix(NA,ncol=7,nrow=numDays)

pool <- list()
for(i in 1:numAgents)
{
   pool[[i]] <- makeAgent(psychstate=1,
                         biostate=sample(c(1,6),
                          p=c(1-naturalImmunity, naturalImmunity),1))
}

##infect patient 0
numInfected <- 5
for(i in sample(numAgents,numInfected))
{
   pool[[i]] <- setAgentState(pool[[i]],2) ##infect this person
}

for(day in 1:numDays)
{

  
    # Create network from individual workplaces/schools/etc. 
  socialnetwork <- net$housenet + net$neighborhood+ net$worknet +net$schoolnet
##who are you going to talk to today.
  if(day>3){
    socialnetwork <- net$housenet+net$neighborhood
  }
    if(day>21){
    socialnetwork <- net$housenet+net$neighborhood+net$worknet
  }
    sneezers <- rep(1:numAgents,each=numInteractions[day])
    sneezedons <- rep(NA,length(sneezers))
    
    for(i in 1:length(sneezers))
      {
      if(runif(1)<(1-sampleFromNetwork[day]) )
      {
        sneezedons[i] <- sample(numAgents,1)
      }else{
        sneezedons[i] <- sample(1:numAgents,prob=socialnetwork[sneezers[i],],1)
      }
     }
      
  for(i in 1:length(sneezers))
  {
    agent1 <- pool[[ sneezers[i] ]]
    agent2 <- pool[[ sneezedons[i] ]]
  
    
    ##this constitutes the rules of infection.
    if((agent1$biostate==2 || agent1$biostate==3 ) & agent2$biostate==1 & runif(1)<contagionProb[day])
    {
         pool[[ sneezedons[i] ]] <- setAgentState(agent2,2)##infect!
    }
      
  }
    ##increment each agent 1-day.
    for(i in 1:numAgents)
    {
      pool[[i]] <- updateAgent(pool[[i]])
    }
    
    states <- sapply(pool,FUN=function(x){x$biostate})
    distrib <- table(factor(states,levels=1:7))
    disthistory[day,] <- distrib
   if(plotNetwork)
   {
     mygplot(cc,socialnetwork,states,main=paste("Day",day))
   }
  
         
}
```

```{r fig.width=8,fig.height=5}
disthist.df <-as.data.frame(disthistory)
colnames(disthist.df) <- STATENAMES
disthist.df$day <- 1:nrow(disthistory)
histlong <- melt(disthist.df,id.vars="day")

ggplot(histlong,aes(x=day,y=value,fill=variable)) + geom_bar(stat="identity",position="stack") + 
  theme_bw()

##make the SIR plot:
sir <- data.frame(day=disthist.df$day,
                  susceptible = disthist.df$Unexposed,
                  infected = disthist.df[,2]+disthist.df[,3],
                  recovered = rowSums(disthist.df[,4:7]))

ggplot(melt(sir,id.vars="day"),aes(x=day,group=variable,y=value,color=variable)) + geom_line() + theme_bw()
ggplot(histlong,aes(x=day,y=value,group=variable,color=variable)) + geom_line() + theme_bw()
```

###Check isolation stage: no school but  no workplace from day 4 to day 27 day
```{r,fig.width=20,fig.height=20}

numAgents <- 500
numDays <- 90
naturalImmunity <- .01
net <- makeDemographicNetwork(numAgents)

pool <- net$pool
cc <- net$xy
socialnetwork <- net$network


numInteractions <-  rep(20,numDays)  ##how many interactions per day per agent on average?
contagionProb <- rep(.05,numDays)    ##normal contagioun probability after concat
sampleFromNetwork <- rep(.99,numDays)  ##how likely you are to stick with 'your' network

numInteractions[4:numDays] <- 5     ##quarantiine goes into effect day 3.
numInteractions[27:numDays] <- 15  ##re-open 

plotNetwork <- TRUE
#re-use previous network
#socialnetwork <-makeNetwork(numAgents,numsets=1,power=.5)


if(plotNetwork)
{
    cc <-mygplot(coord=cc,socialnetwork,rep(1,nrow(socialnetwork)),main="Initial state")

}



disthistory <- matrix(NA,ncol=7,nrow=numDays)

pool <- list()
for(i in 1:numAgents)
{
   pool[[i]] <- makeAgent(psychstate=1,
                         biostate=sample(c(1,6),
                          p=c(1-naturalImmunity, naturalImmunity),1))
}

##infect patient 0
numInfected <- 5
for(i in sample(numAgents,numInfected))
{
   pool[[i]] <- setAgentState(pool[[i]],2) ##infect this person
}

for(day in 1:numDays)
{

  # Create network from individual workplaces/schools/etc. 
   socialnetwork <- net$housenet + net$neighborhood+ net$worknet +net$schoolnet
##who are you going to talk to today.
  if(day>3){
    socialnetwork <- net$housenet+net$neighborhood
  }
  if(day>27){
    socialnetwork <- net$housenet+net$neighborhood +net$worknet
  }
    sneezers <- rep(1:numAgents,each=numInteractions[day])
    sneezedons <- rep(NA,length(sneezers))
    
    for(i in 1:length(sneezers))
      {
      if(runif(1)<(1-sampleFromNetwork[day]) )
      {
        sneezedons[i] <- sample(numAgents,1)
      }else{
        sneezedons[i] <- sample(1:numAgents,prob=socialnetwork[sneezers[i],],1)
      }
     }
      
  for(i in 1:length(sneezers))
  {
    agent1 <- pool[[ sneezers[i] ]]
    agent2 <- pool[[ sneezedons[i] ]]
  
    
    ##this constitutes the rules of infection.
    if((agent1$biostate==2 || agent1$biostate==3 ) & agent2$biostate==1 & runif(1)<contagionProb[day])
    {
         pool[[ sneezedons[i] ]] <- setAgentState(agent2,2)##infect!
    }
      
  }
    ##increment each agent 1-day.
    for(i in 1:numAgents)
    {
      pool[[i]] <- updateAgent(pool[[i]])
    }
    
    states <- sapply(pool,FUN=function(x){x$biostate})
    distrib <- table(factor(states,levels=1:7))
    disthistory[day,] <- distrib
   if(plotNetwork)
   {
     mygplot(cc,socialnetwork,states,main=paste("Day",day))
   }
  
         
}
```
```{r fig.width=8,fig.height=5}
disthist.df <-as.data.frame(disthistory)
colnames(disthist.df) <- STATENAMES
disthist.df$day <- 1:nrow(disthistory)
histlong <- melt(disthist.df,id.vars="day")

ggplot(histlong,aes(x=day,y=value,fill=variable)) + geom_bar(stat="identity",position="stack") + 
  theme_bw()

##make the SIR plot:
sir <- data.frame(day=disthist.df$day,
                  susceptible = disthist.df$Unexposed,
                  infected = disthist.df[,2]+disthist.df[,3],
                  recovered = rowSums(disthist.df[,4:7]))

ggplot(melt(sir,id.vars="day"),aes(x=day,group=variable,y=value,color=variable)) + geom_line() + theme_bw()
ggplot(histlong,aes(x=day,y=value,group=variable,color=variable)) + geom_line() + theme_bw()
```





###Check isolation stage: no school but  no workplace from day 4 to day 34 
```{r,fig.width=20,fig.height=20}

numAgents <- 500
numDays <- 90
naturalImmunity <- .01
net <- makeDemographicNetwork(numAgents)

pool <- net$pool
cc <- net$xy
socialnetwork <- net$network


numInteractions <-  rep(20,numDays)  ##how many interactions per day per agent on average?
contagionProb <- rep(.05,numDays)    ##normal contagioun probability after concat
sampleFromNetwork <- rep(.99,numDays)  ##how likely you are to stick with 'your' network

numInteractions[4:numDays] <- 5     ##quarantiine goes into effect day 3.
numInteractions[34:numDays] <- 15  ##re-open 

plotNetwork <- TRUE
#re-use previous network
#socialnetwork <-makeNetwork(numAgents,numsets=1,power=.5)


if(plotNetwork)
{
    cc <-mygplot(coord=cc,socialnetwork,rep(1,nrow(socialnetwork)),main="Initial state")

}



disthistory <- matrix(NA,ncol=7,nrow=numDays)

pool <- list()
for(i in 1:numAgents)
{
   pool[[i]] <- makeAgent(psychstate=1,
                         biostate=sample(c(1,6),
                          p=c(1-naturalImmunity, naturalImmunity),1))
}

##infect patient 0
numInfected <- 5
for(i in sample(numAgents,numInfected))
{
   pool[[i]] <- setAgentState(pool[[i]],2) ##infect this person
}

for(day in 1:numDays)
{

  # Create network from individual workplaces/schools/etc. 
  socialnetwork <- net$housenet + net$neighborhood+ net$worknet +net$schoolnet
##who are you going to talk to today.
  if(day>3){
    socialnetwork <- net$housenet+net$neighborhood
  }
  if(day>34){
    socialnetwork <- net$housenet+net$neighborhood +net$worknet
  }
    sneezers <- rep(1:numAgents,each=numInteractions[day])
    sneezedons <- rep(NA,length(sneezers))
    
    for(i in 1:length(sneezers))
      {
      if(runif(1)<(1-sampleFromNetwork[day]) )
      {
        sneezedons[i] <- sample(numAgents,1)
      }else{
        sneezedons[i] <- sample(1:numAgents,prob=socialnetwork[sneezers[i],],1)
      }
     }
      
  for(i in 1:length(sneezers))
  {
    agent1 <- pool[[ sneezers[i] ]]
    agent2 <- pool[[ sneezedons[i] ]]
  
    
    ##this constitutes the rules of infection.
    if((agent1$biostate==2 || agent1$biostate==3 ) & agent2$biostate==1 & runif(1)<contagionProb[day])
    {
         pool[[ sneezedons[i] ]] <- setAgentState(agent2,2)##infect!
    }
      
  }
    ##increment each agent 1-day.
    for(i in 1:numAgents)
    {
      pool[[i]] <- updateAgent(pool[[i]])
    }
    
    states <- sapply(pool,FUN=function(x){x$biostate})
    distrib <- table(factor(states,levels=1:7))
    disthistory[day,] <- distrib
   if(plotNetwork)
   {
     mygplot(cc,socialnetwork,states,main=paste("Day",day))
   }
  
         
}
```
```{r fig.width=8,fig.height=5}
disthist.df <-as.data.frame(disthistory)
colnames(disthist.df) <- STATENAMES
disthist.df$day <- 1:nrow(disthistory)
histlong <- melt(disthist.df,id.vars="day")

ggplot(histlong,aes(x=day,y=value,fill=variable)) + geom_bar(stat="identity",position="stack") + 
  theme_bw()

##make the SIR plot:
sir <- data.frame(day=disthist.df$day,
                  susceptible = disthist.df$Unexposed,
                  infected = disthist.df[,2]+disthist.df[,3],
                  recovered = rowSums(disthist.df[,4:7]))

ggplot(melt(sir,id.vars="day"),aes(x=day,group=variable,y=value,color=variable)) + geom_line() + theme_bw()
ggplot(histlong,aes(x=day,y=value,group=variable,color=variable)) + geom_line() + theme_bw()
```

###Check isolation stage: no school but  no workplace from day 4 to day 41 
```{r,fig.width=20,fig.height=20}

numAgents <- 500
numDays <- 90
naturalImmunity <- .01
net <- makeDemographicNetwork(numAgents)

pool <- net$pool
cc <- net$xy
socialnetwork <- net$network


numInteractions <-  rep(20,numDays)  ##how many interactions per day per agent on average?
contagionProb <- rep(.05,numDays)    ##normal contagioun probability after concat
sampleFromNetwork <- rep(.99,numDays)  ##how likely you are to stick with 'your' network

numInteractions[4:numDays] <- 5     ##quarantiine goes into effect day 3.
numInteractions[41:numDays] <- 15  ##re-open 

plotNetwork <- TRUE
#re-use previous network
#socialnetwork <-makeNetwork(numAgents,numsets=1,power=.5)


if(plotNetwork)
{
    cc <-mygplot(coord=cc,socialnetwork,rep(1,nrow(socialnetwork)),main="Initial state")

}



disthistory <- matrix(NA,ncol=7,nrow=numDays)

pool <- list()
for(i in 1:numAgents)
{
   pool[[i]] <- makeAgent(psychstate=1,
                         biostate=sample(c(1,6),
                          p=c(1-naturalImmunity, naturalImmunity),1))
}

##infect patient 0
numInfected <- 5
for(i in sample(numAgents,numInfected))
{
   pool[[i]] <- setAgentState(pool[[i]],2) ##infect this person
}

for(day in 1:numDays)
{

  # Create network from individual workplaces/schools/etc. 
  socialnetwork <- net$housenet + net$neighborhood+ net$worknet +net$schoolnet
##who are you going to talk to today.
  if(day>3){
    socialnetwork <- net$housenet+net$neighborhood
  }
  if(day>41){
    socialnetwork <- net$housenet+net$neighborhood +net$worknet
  }
    sneezers <- rep(1:numAgents,each=numInteractions[day])
    sneezedons <- rep(NA,length(sneezers))
    
    for(i in 1:length(sneezers))
      {
      if(runif(1)<(1-sampleFromNetwork[day]) )
      {
        sneezedons[i] <- sample(numAgents,1)
      }else{
        sneezedons[i] <- sample(1:numAgents,prob=socialnetwork[sneezers[i],],1)
      }
     }
      
  for(i in 1:length(sneezers))
  {
    agent1 <- pool[[ sneezers[i] ]]
    agent2 <- pool[[ sneezedons[i] ]]
  
    
    ##this constitutes the rules of infection.
    if((agent1$biostate==2 || agent1$biostate==3 ) & agent2$biostate==1 & runif(1)<contagionProb[day])
    {
         pool[[ sneezedons[i] ]] <- setAgentState(agent2,2)##infect!
    }
      
  }
    ##increment each agent 1-day.
    for(i in 1:numAgents)
    {
      pool[[i]] <- updateAgent(pool[[i]])
    }
    
    states <- sapply(pool,FUN=function(x){x$biostate})
    distrib <- table(factor(states,levels=1:7))
    disthistory[day,] <- distrib
   if(plotNetwork)
   {
     mygplot(cc,socialnetwork,states,main=paste("Day",day))
   }
  
         
}
```

```{r fig.width=8,fig.height=5}
disthist.df <-as.data.frame(disthistory)
colnames(disthist.df) <- STATENAMES
disthist.df$day <- 1:nrow(disthistory)
histlong <- melt(disthist.df,id.vars="day")

ggplot(histlong,aes(x=day,y=value,fill=variable)) + geom_bar(stat="identity",position="stack") + 
  theme_bw()

##make the SIR plot:
sir <- data.frame(day=disthist.df$day,
                  susceptible = disthist.df$Unexposed,
                  infected = disthist.df[,2]+disthist.df[,3],
                  recovered = rowSums(disthist.df[,4:7]))

ggplot(melt(sir,id.vars="day"),aes(x=day,group=variable,y=value,color=variable)) + geom_line() + theme_bw()
ggplot(histlong,aes(x=day,y=value,group=variable,color=variable)) + geom_line() + theme_bw()
```



###Check isolation stage: no school but  no workplace from day 4 to day 48 
```{r,fig.width=20,fig.height=20}

numAgents <- 500
numDays <- 90
naturalImmunity <- .01
net <- makeDemographicNetwork(numAgents)

pool <- net$pool
cc <- net$xy
socialnetwork <- net$network


numInteractions <-  rep(20,numDays)  ##how many interactions per day per agent on average?
contagionProb <- rep(.05,numDays)    ##normal contagioun probability after concat
sampleFromNetwork <- rep(.99,numDays)  ##how likely you are to stick with 'your' network

numInteractions[4:numDays] <- 5     ##quarantiine goes into effect day 3.
numInteractions[48:numDays] <- 15  ##re-open 

plotNetwork <- TRUE
#re-use previous network
#socialnetwork <-makeNetwork(numAgents,numsets=1,power=.5)


if(plotNetwork)
{
    cc <-mygplot(coord=cc,socialnetwork,rep(1,nrow(socialnetwork)),main="Initial state")

}



disthistory <- matrix(NA,ncol=7,nrow=numDays)

pool <- list()
for(i in 1:numAgents)
{
   pool[[i]] <- makeAgent(psychstate=1,
                         biostate=sample(c(1,6),
                          p=c(1-naturalImmunity, naturalImmunity),1))
}

##infect patient 0
numInfected <- 5
for(i in sample(numAgents,numInfected))
{
   pool[[i]] <- setAgentState(pool[[i]],2) ##infect this person
}

for(day in 1:numDays)
{

  # Create network from individual workplaces/schools/etc. 
  socialnetwork <- net$housenet + net$neighborhood+ net$worknet +net$schoolnet
##who are you going to talk to today.
  if(day>3){
    socialnetwork <- net$housenet+net$neighborhood
  }
  if(day>48){
    socialnetwork <- net$housenet+net$neighborhood +net$worknet
  }
    sneezers <- rep(1:numAgents,each=numInteractions[day])
    sneezedons <- rep(NA,length(sneezers))
    
    for(i in 1:length(sneezers))
      {
      if(runif(1)<(1-sampleFromNetwork[day]) )
      {
        sneezedons[i] <- sample(numAgents,1)
      }else{
        sneezedons[i] <- sample(1:numAgents,prob=socialnetwork[sneezers[i],],1)
      }
     }
      
  for(i in 1:length(sneezers))
  {
    agent1 <- pool[[ sneezers[i] ]]
    agent2 <- pool[[ sneezedons[i] ]]
  
    
    ##this constitutes the rules of infection.
    if((agent1$biostate==2 || agent1$biostate==3 ) & agent2$biostate==1 & runif(1)<contagionProb[day])
    {
         pool[[ sneezedons[i] ]] <- setAgentState(agent2,2)##infect!
    }
      
  }
    ##increment each agent 1-day.
    for(i in 1:numAgents)
    {
      pool[[i]] <- updateAgent(pool[[i]])
    }
    
    states <- sapply(pool,FUN=function(x){x$biostate})
    distrib <- table(factor(states,levels=1:7))
    disthistory[day,] <- distrib
   if(plotNetwork)
   {
     mygplot(cc,socialnetwork,states,main=paste("Day",day))
   }
  
         
}
```

```{r fig.width=8,fig.height=5}
disthist.df <-as.data.frame(disthistory)
colnames(disthist.df) <- STATENAMES
disthist.df$day <- 1:nrow(disthistory)
histlong <- melt(disthist.df,id.vars="day")

ggplot(histlong,aes(x=day,y=value,fill=variable)) + geom_bar(stat="identity",position="stack") + 
  theme_bw()

##make the SIR plot:
sir <- data.frame(day=disthist.df$day,
                  susceptible = disthist.df$Unexposed,
                  infected = disthist.df[,2]+disthist.df[,3],
                  recovered = rowSums(disthist.df[,4:7]))

ggplot(melt(sir,id.vars="day"),aes(x=day,group=variable,y=value,color=variable)) + geom_line() + theme_bw()
ggplot(histlong,aes(x=day,y=value,group=variable,color=variable)) + geom_line() + theme_bw()
```


###Check isolation stage: no school but  no workplace from day 4 to day 55 
```{r,fig.width=20,fig.height=20}

numAgents <- 500
numDays <- 90
naturalImmunity <- .01
net <- makeDemographicNetwork(numAgents)

pool <- net$pool
cc <- net$xy
socialnetwork <- net$network


numInteractions <-  rep(20,numDays)  ##how many interactions per day per agent on average?
contagionProb <- rep(.05,numDays)    ##normal contagioun probability after concat
sampleFromNetwork <- rep(.99,numDays)  ##how likely you are to stick with 'your' network

numInteractions[4:numDays] <- 5     ##quarantiine goes into effect day 3.
numInteractions[55:numDays] <- 15  ##re-open 

plotNetwork <- TRUE
#re-use previous network
#socialnetwork <-makeNetwork(numAgents,numsets=1,power=.5)


if(plotNetwork)
{
    cc <-mygplot(coord=cc,socialnetwork,rep(1,nrow(socialnetwork)),main="Initial state")

}



disthistory <- matrix(NA,ncol=7,nrow=numDays)

pool <- list()
for(i in 1:numAgents)
{
   pool[[i]] <- makeAgent(psychstate=1,
                         biostate=sample(c(1,6),
                          p=c(1-naturalImmunity, naturalImmunity),1))
}

##infect patient 0
numInfected <- 5
for(i in sample(numAgents,numInfected))
{
   pool[[i]] <- setAgentState(pool[[i]],2) ##infect this person
}

for(day in 1:numDays)
{

  # Create network from individual workplaces/schools/etc. 
  socialnetwork <- net$housenet + net$neighborhood+ net$worknet +net$schoolnet
##who are you going to talk to today.
  if(day>3){
    socialnetwork <- net$housenet+net$neighborhood
  }
  if(day>55){
    socialnetwork <- net$housenet+net$neighborhood +net$worknet
  }
    sneezers <- rep(1:numAgents,each=numInteractions[day])
    sneezedons <- rep(NA,length(sneezers))
    
    for(i in 1:length(sneezers))
      {
      if(runif(1)<(1-sampleFromNetwork[day]) )
      {
        sneezedons[i] <- sample(numAgents,1)
      }else{
        sneezedons[i] <- sample(1:numAgents,prob=socialnetwork[sneezers[i],],1)
      }
     }
      
  for(i in 1:length(sneezers))
  {
    agent1 <- pool[[ sneezers[i] ]]
    agent2 <- pool[[ sneezedons[i] ]]
  
    
    ##this constitutes the rules of infection.
    if((agent1$biostate==2 || agent1$biostate==3 ) & agent2$biostate==1 & runif(1)<contagionProb[day])
    {
         pool[[ sneezedons[i] ]] <- setAgentState(agent2,2)##infect!
    }
      
  }
    ##increment each agent 1-day.
    for(i in 1:numAgents)
    {
      pool[[i]] <- updateAgent(pool[[i]])
    }
    
    states <- sapply(pool,FUN=function(x){x$biostate})
    distrib <- table(factor(states,levels=1:7))
    disthistory[day,] <- distrib
   if(plotNetwork)
   {
     mygplot(cc,socialnetwork,states,main=paste("Day",day))
   }
  
         
}
```

```{r fig.width=8,fig.height=5}
disthist.df <-as.data.frame(disthistory)
colnames(disthist.df) <- STATENAMES
disthist.df$day <- 1:nrow(disthistory)
histlong <- melt(disthist.df,id.vars="day")

ggplot(histlong,aes(x=day,y=value,fill=variable)) + geom_bar(stat="identity",position="stack") + 
  theme_bw()

##make the SIR plot:
sir <- data.frame(day=disthist.df$day,
                  susceptible = disthist.df$Unexposed,
                  infected = disthist.df[,2]+disthist.df[,3],
                  recovered = rowSums(disthist.df[,4:7]))

ggplot(melt(sir,id.vars="day"),aes(x=day,group=variable,y=value,color=variable)) + geom_line() + theme_bw()
ggplot(histlong,aes(x=day,y=value,group=variable,color=variable)) + geom_line() + theme_bw()
```


