---
title: "Modeling disease spread with a geographic agent-based model"
author: "Anne Linja and Shane Mueller"
date: "May 5, 2020"
output:
  rmdformats::readthedown:
    gallery: yes
    highlight: kate
    self_contained: no
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---
This simulation is an adaptation of the base model described [epidemic-base.html](here).  In this simulation, we generate a network model based roughly on the geographic layout of michigan, in order to explore how geography may impact disease spread.
 
## The agent
 

 The biological state has a few specific cases:
 

1. Unexposed
2. Asymptomatic but infected/contagious
3. Symptomatic and contagious
4. Symptomatic and not contagious
5. Post-COVID Immune
6. Naturally immune (will not contract)
7. Death


We could identify several others, but this might be fine.  So, we can define the agent according to just two numbers.  We might assume that initially, most people are in bio-state 1,  but some would be in state 6 already, which is essentially the same as state 5.

We will just define these as a named list.
```{r}
library(ggplot2)
library(dplyr)
library(sna)
library(knitr)
library(reshape2)
library(igraph)
#library(animation)
STATES <<- 7

STATENAMES <<-  c("Unexposed",
                  "Asymptomatic & contagious",
"Symptomatic and contagious",
"Symptomatic and not contagious",
"Post-COVID immune",
"Naturally immune",
"Death")


STATELABELS <<-  c("Unexposed","Asymptomatic\n & contagious",
"Symptomatic \n& contagious",
"Symptomatic \n& not contagious",
"Post-COVID immune",
"Naturally immune",
"Death")

makeAgent <- function(psychstate,biostate,age=30)
{
  
  return (list(psychstate=psychstate,
               biostate=biostate,
               age=age,
               nextbiostate=NA,
               biostatecountdown=NA))
}
print(makeAgent(1,2))


updateAgent<- function(agent)
{
   agent$biostatecountdown <- agent$biostatecountdown -1
  if(agent$biostatecountdown <=0)
  {
    agent$biostate <- agent$nextbiostate
    agent$biostatecountdown <- NA
  }
   return(agent)
}
```




## Create transition matrix.

If on each timestep we could move between states probabilistically, this would be a markov process. We could represent this by a simple transition network. If we represented the possible transitions in a network, that would simplify the logic above.  But we need more like a task network because the time taken to transition matters.   To keep it simple,   Let's make all timing distributions uniform with a min and max parameter for time in each state.  We can program two pathways through the stages, with a couple branch points (death vs recovery; the possibility of recovering after acquiring with no symptoms).  The progression of the disease is completely specified by the data, and a generic update function will automatically progress the agent each day.


```{r}

# * 1. Unexposed
# * 2. Asymptomatic but infected/contagious
# * 3. Symptomatic and contagious
# * 4. Symptomatic and not contagious
# * 5. Post-COVID Immune
# * 6. Naturally immune (will not contract)
# * 7. Death


bioTransition <- matrix(0,STATES,STATES)
bioMin <- matrix(1,STATES)      #state time minimum
bioMax <- matrix(1,STATES)      #state time maximum



bioMin[2] <- 3            #infected but asymptomatic for 3 to 10 days
bioMax[2] <- 10          
bioTransition[2,3] <- .5  #transition to infected with symptoms
bioTransition[2,5] <- .5  #transition to no longer contagious/cured


bioMin[3] <-    3             #symptoms + contagion
bioMax[3] <- 8                #symptoms + contagion max
bioTransition[3,4] <-  .95    #transitioon to no longer contagious
bioTransition[3,7] <-  .05    #transitioon to death state 


bioMin[4] <- 1          #symptoms bot no longer contagiious
bioMax[4] <- 7
bioTransition[4,5] <- 1  #Transition to 'immune' cured state.




setAgentState<- function(agent, biostate)
{
  agent$biostate <- biostate
  if(sum(bioTransition[biostate,])>0) # this state transitions to something else.
  {
    ##which state do we go to?
     agent$biostatecountdown <- sample(x=seq(bioMin[biostate],bioMax[biostate]),1) #how long will we state in this state?
     agent$nextbiostate <- sample(1:STATES, prob=bioTransition[agent$biostate,],size=1)
     
  } else{
   agent$biostatecountdown <- NA
   agent$nextbiostate <- NA   ##just so we can tell if the agent is finished.
  }
  return(agent) 
}

transitionAgent<- function(agent)
{
   return(setAgentState(agent,agent$nextbiostate))
}

updateAgent<- function(agent)
{
  if(!is.na(agent$biostatecountdown))
  {
   agent$biostatecountdown <- agent$biostatecountdown -1
    if(agent$biostatecountdown <=0)  ##new state
    {
       agent <- transitionAgent(agent)

    }
  }
   return(agent)
}
```

```{r, fig.width=8,fig.height=8}

par(xpd=NA,mar=c(5,5,8,5))
gplot(bioTransition,label=STATELABELS)
```


```{r}

mygplot <- function(coord, network,states,main="")
{
  if(is.null(coord))
  {
    coord  <- gplot.layout.fruchtermanreingold(network,layout.par=list(niter=500))
  }
  
  newmin <- mean(coord[,2]) -  (-min(coord[,2]) + mean(coord[,2])) * 1.4
   palette=c("white","yellow","red","green","darkgreen","blue","black")
   plot(coord,col="black",bty="n",pch=16,cex=2.7,xaxt="n",yaxt="n",main=main,xlab="",
          ylab="",axes=F,
        ylim=c(newmin,max(coord[,2])),type="n")
   
  for(i in 1:nrow(network))
     segments(coord[i,1],
            coord[i,2],
            coord[network[i,]==1,1],
            coord[network[i,]==1,2],col="grey40")

       points(coord,pch=16,cex=2.3,col= palette[states])
              
       points(coord,pch=1,cex=2.3,col="black")
      legend(mean(coord[,1]),min(coord[,2]),bty='n',y.intersp=.7,cex=.8,
              STATENAMES, pch=16,col=palette)
              
      return (coord)
  }

```

## Modeling michigan county data.

```{r,fig.width=18,fig.height=20}
#library(readxl)
counties   <- read.csv("MI_Counties.csv")
numcounties <- nrow(counties) ##83

plot(counties$LON,counties$LAT,pch=16,cex=3)

countydist <- as.matrix(dist(cbind(counties$LON,counties$LAT*2)))
ranks <- apply(countydist,1,rank)
nearby <- 0+(ranks<7 | countydist<.85 )



coords <- data.frame(LON=counties$LON,LAT=counties$LAT*2)
coords$UP <- coords$LAT>91.05

mygplot(coord=coords[,1:2],nearby,states=coords$UP+1)
#matrix(NA,nrow=nrow(countydist),ncol=nrow(countydist))

numAgents <- 2500 
simPop <- ceiling(counties$Population/sum(counties$Population) * numAgents) 
numAgents <- sum(simPop)
whichcounty <- rep(1:numcounties,times=simPop) 
inUP <- coords$UP[whichcounty]


##Create the network

withincounty <- outer(whichcounty,whichcounty,"==") 
betweencounties <- matrix(NA,numAgents, numAgents) 
for(i in 1:numAgents)
{ 
  for(j in 1:numAgents)
   {
    betweencounties[i,j] <-runif(1) < .5* nearby[whichcounty[i],whichcounty[j]]
   }
}
#gplot(withincounty+betweencounties)
socialnetwork <- (withincounty + betweencounties) > 0 #this is where we make the whole network. either T or 1, or F/2.
coords2 <- cbind(coords[whichcounty,1] + rnorm(numAgents)*.15,
                    coords[whichcounty,2] + rnorm(numAgents)*.15) 


if(TRUE) 
{
    cc <-mygplot(coord=coords2,socialnetwork,rep(1,nrow(socialnetwork)),main="Initial state")

}
```

##Now, make the social network combine within and between county data. Simulate disease spread.

```{r,fig.width=10,fig.height=12, fig.show='animate'}

#mygplot(coord=coords2,socialnetwork,
#        states=rep(1,numAgents))


numDays <- 56

naturalImmunity <- .01  


numInteractions <-  rep(10,numDays)  ##how many interactions per day per agent on average initially?
contagionProb <- rep(.05,numDays)    ##normal contagian probability after contact
sampleFromNetwork <- rep(1.0,numDays)  ##how likely you are to stick with 'your' network

numInteractions[5:numDays] <- 5     
numInteractions[25:numDays] <- 10  

plotNetwork <- TRUE
#re-use previous network
#socialnetwork <-makeNetwork(numAgents,numsets=1,power=.5)


if(plotNetwork)
{
    cc <-mygplot(coord=coords2,socialnetwork,rep(1,nrow(socialnetwork)),main="Initial state")

}


disthistory <- matrix(NA,ncol=7,nrow=numDays)
disthistory.up <- matrix(NA,ncol=7,nrow=numDays)

pool <- list()
for(i in 1:numAgents)
{
   pool[[i]] <- makeAgent(psychstate=1,
                         biostate=sample(c(1,6),
                          p=c(1-naturalImmunity, naturalImmunity),1))
}


##Here we can decide who gets infected first:
##infect patient 0
numInfected <- 5 
for(i in sample(numAgents,numInfected))
{
   pool[[i]] <- setAgentState(pool[[i]],2) ##infect this person
}



for(day in 1:numDays)
{
##who are you going to talk to today.
    sneezers <- rep(1:numAgents,each=numInteractions[day])
    sneezedons <- rep(NA,length(sneezers))
    
    for(i in 1:length(sneezers))
      {
      if(runif(1)<(1-sampleFromNetwork[day]) )
      {
        sneezedons[i] <- sample(numAgents,1)
      }else{
        sneezedons[i] <- sample(1:numAgents,prob=socialnetwork[sneezers[i],],1)
      }
     }
      
  for(i in 1:length(sneezers))
  {
    agent1 <- pool[[ sneezers[i] ]]
    agent2 <- pool[[ sneezedons[i] ]]
  
    
    ##this constitutes the rules of infection.
    if((agent1$biostate==2 || agent1$biostate==3 ) & agent2$biostate==1 & runif(1)<contagionProb[day])
    {
         pool[[ sneezedons[i] ]] <- setAgentState(agent2,2)##infect!
    }
      
  }
    ##increment each agent 1-day.
    for(i in 1:numAgents)
    {
      pool[[i]] <- updateAgent(pool[[i]])
    }
    
    states <- sapply(pool,FUN=function(x){x$biostate})
    states.up <- sapply(pool[inUP],FUN=function(x){x$biostate})
    
    distrib <- table(factor(states,levels=1:7))
    distrib.up <- table(factor(states.up,levels=1:7))
    
    disthistory[day,] <- distrib
    disthistory.up[day,] <- distrib.up

       if(plotNetwork)
   {
     mygplot(cc,socialnetwork,states,main=paste("Day",day))
   }
  
         
  }
  
  #barplot(t(disthistory),col=1:7)
  
```
## Plot final state
```{r,fig.width=15,fig.height=20}
 xx <- mygplot(cc,socialnetwork,states,main=paste("Day",day))
```

## Progress in state vs. UP only. Growth for UP and for entire state. 

```{r fig.width=8,fig.height=5} 
disthist.df <-as.data.frame(disthistory)
colnames(disthist.df) <- STATENAMES
disthist.df$day <- 1:nrow(disthistory)
histlong <- melt(disthist.df,id.vars="day")

ggplot(histlong,aes(x=day,y=value,fill=variable)) + geom_bar(stat="identity",position="stack") + 
  theme_bw()

##make the SIR plot:
sir <- data.frame(day=disthist.df$day,
                  susceptible = disthist.df$Unexposed,
                  infected = disthist.df[,2]+disthist.df[,3],
                  recovered = rowSums(disthist.df[,4:7]))

ggplot(melt(sir,id.vars="day"),aes(x=day,group=variable,y=value,color=variable)) + geom_line() + theme_bw()+ ggtitle(label="Disease spread in Michigan")

ggplot(histlong,aes(x=day,y=value,group=variable,color=variable)) + geom_line() + theme_bw() + ggtitle(label="Disease spread in Michigan")


##Now, do UP only

disthist.df <-as.data.frame(disthistory.up)
colnames(disthist.df) <- STATENAMES
disthist.df$day <- 1:nrow(disthistory.up)
histlong <- melt(disthist.df,id.vars="day")

ggplot(histlong,aes(x=day,y=value,fill=variable)) + geom_bar(stat="identity",position="stack") + 
  theme_bw() +  ggtitle(label="Disease spread in Upper Peninsula Michigan")



##make the SIR plot:
sir <- data.frame(day=disthist.df$day,
                  susceptible = disthist.df$Unexposed,
                  infected = disthist.df[,2]+disthist.df[,3],
                  recovered = rowSums(disthist.df[,4:7]))


ggplot(melt(sir,id.vars="day"),aes(x=day,group=variable,y=value,color=variable)) + geom_line() + theme_bw() + ggtitle(label="Disease spread in Upper Peninsula Michigan")

ggplot(histlong,aes(x=day,y=value,group=variable,color=variable)) + geom_line() + theme_bw()+ ggtitle(label="Disease spread in Upper Peninsula Michigan")

```

##can model each county as an SIR. Can model larger numagents, but it takes forever to run.